---
title: "R4GIS Notebook"
output:
  html_document:
    df_print: paged
date: "`r Sys.time()`"
---

# About
This R notebook is intended to be the R code of the course Automating GIS-processes (https://autogis-site.readthedocs.io/en/latest/index.html). The data are available in the web site of the course.
The learning objectives of the course are:

* read / write spatial data from/to different file formats
* work with coordinate reference systems
* geocode data from addresses to coordinates and vice versa
* apply geometric operations (intersect, union)
* reclassify data based on conditions
* apply spatial queries
* conduct simple spatial analyses
* Visualize data and create (interactive) maps
```{r libraries, message=FALSE, warning=FALSE, echo=FALSE}
suppressWarnings(suppressPackageStartupMessages(library(tidyverse)))
suppressPackageStartupMessages(library(sf))
suppressPackageStartupMessages(library(tmap))
suppressPackageStartupMessages(library(httr))
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(library(ows4R))
#suppressPackageStartupMessages(library(arcgisbinding))
```
# 0. environment set up
Install required packages
```{r}
libs <- c("sf", "tmap", "sp", "spdep", "DCluster", "cartogram")
for (x in libs){                      #cycle through each item in libs object
  if(x %in% rownames(installed.packages()) == FALSE) {  #if the package is not installed
    print(paste0("installing ", x, "...")) #print a message to tell me
    install.packages(x)                 #and then install the packages
  }
  else{                                     #otherwise (if it is installed)
    print (paste0(x, " is already installed "))     #print a message to tell me
  }
 
  library(x, character.only = TRUE)               #and then load the packages
 
}

```
 

# Lesson 1: Spatial data model
 
**Learning goals:** Create and manipulate Geometric objects
 
* Points, Linestring and Polygon in R
 

```{r}
point1 <- sf::st_point(x = c(2.2,4.2),dim = "XY")
point2 <- sf::st_point(x = c(7.2, -25.1),dim = "XY")
point3 <- sf::st_point(x = c(9.26, -2.456),dim = "XY")
point3D <- sf::st_point(x = c(9.26, -2.456, 0.57),dim = "XYZ")
line <- sf::st_linestring(rbind(point1,point2,point3))
line2 <- rbind(c(2.2, 4.2), c(7.2, -25.1), c(9.26, -2.456))
line2 <- sf::st_linestring(line2)
poly <- sf::st_polygon(list(rbind(c(2.2, 4.2), c(7.2, -25.1), c(9.26, -2.456),c(2.2, 4.2))),dim = "XY")
poly2 <- sf::st_polygon(list(rbind(c(-180, 90), c(-180, -90), c(180, -90), c(180, 90),c(-180, 90)),rbind(c(-170, 80), c(-170, -80), c(170, -80), c(170, 80),c(-170, 80))  ))
world_exterior <- rbind(c(-180, 90), c(-180, -90), c(180, -90), c(180, 90),c(-180, 90))
hole <- rbind(c(-170, 80), c(-170, -80), c(170, -80), c(170, 80),c(-170, 80))
world <- sf::st_polygon(list(world_exterior,hole))
```
 

```{r}
par(mfrow=c(1,4))
ggplot() + geom_sf(data = point1) + ggtitle("Point")
ggplot() + geom_sf(data = line) + ggtitle("Linestring")
ggplot() + geom_sf(data = poly) + ggtitle("Polygon")
ggplot() + geom_sf(data=world) + ggtitle("Outer polygon with hole")
```
 

* Geometry collection
 
```{r}
multi_point <- sf::st_multipoint(x = c(point1,point2,point3))
line1 <- sf::st_linestring(c(point1,point2))
line2 <- sf::st_linestring(c(point2,point3))
multi_line <- sf::st_multilinestring(x = list(line1,line2))
```

```{r geom_collection}
par(mfrow=c(1,2))
ggplot()+geom_sf(data=multi_point) + ggtitle("Multipoint")
ggplot()+geom_sf(data=multi_line) + ggtitle("Multilinestring")
```

* Convex hull
 
```{r}
ggplot()+geom_sf(data=sf::st_convex_hull(multi_point))
```

# Lesson 2
 
**Learning goals:**

* Read and write spatial data from/to common file formats
* Conduct simple analysis on spatial and non-spatial data
* Manage coordinate reference systems and re-project data

## Read / write Shapefile
```{r}
shapedata <- sf::st_read("./Automating-GIS-processes/source/notebooks/L2/L2_data/NLS/2018/L4/L41/L4132R.shp/m_L4132R_p.shp")
```
Let’s check the first rows of data and column names:
```{r}
shapedata %>% dplyr::glimpse()
```
 

The column names are in Finnish then Let’s select only the useful columns and rename them into English:

```{r}
shapedata <- shapedata %>% dplyr::rename(GROUP=RYHMA, CLASS=LUOKKA) %>% dplyr::select(GROUP,CLASS)
```
Let's check:

```{r}
shapedata %>% glimpse()
```
Let’s try it out, and plot our GeoDataFrame:
```{r}
ggplot() + geom_sf(data = shapedata)
```
 
```{r}
shapedata %>% head()
```
 
Get information about the area and create a new column that stores the areas of individual polygons:

```{r}
shapedata$area <- shapedata %>% sf::st_area() %>% round()
```

## Writing data into a shapefile

Let’s first select one class (class number 36200, “Lake water”) from the data
 
```{r}
selection <- shapedata %>% dplyr::filter(CLASS==36200)
```
Check the selection:
```{r}
selection %>% ggplot()+geom_sf()
```
 
* Supported Drivers
 
```{r}
sf::st_drivers()
```

```{r}
sf::st_write(obj = selection, dsn="Class_36200.shp")
```

```{r}
basename <- "terrain"
output_folder <- "output/"
if (!dir.exists("output/")){
  dir.create("output/")
}
for(class in unique(shapedata$CLASS)){
  output_name = paste(output_folder,basename,"_",class,".shp",sep="")
  # cat("saving file:", output_name)
  sf::st_write(obj = shapedata %>% filter(CLASS==class), dsn=output_name)
  }
```

## Map projections
 
```{r}
data <- sf::st_read("./Automating-GIS-processes/source/notebooks/L2/L2_data/Europe_borders.shp")
```
 

```{r}
latitudes <- c(63.748023986816406, 62.90789794921875, 60.511383056640625, 60.44499588012695, 60.646385192871094, 60.243743896484375, 59.806800842285156, 59.91944122314453, 60.02395248413086, 60.14555358886719, 60.3452033996582, 60.211936950683594, 60.56249237060547, 61.54027557373047, 62.59798049926758, 63.02013397216797, 63.20353698730469, 63.27652359008789, 63.525691986083984, 64.79915618896484, 64.9533920288086, 65.51513671875, 65.65470886230469, 65.89610290527344, 65.79151916503906, 66.26332092285156, 66.80228424072266, 67.1570053100586, 67.4168701171875, 67.47978210449219, 67.94589233398438, 69.060302734375, 69.32611083984375, 68.71110534667969, 68.83248901367188, 68.580810546875, 68.98916625976562, 69.68568420410156, 69.9363784790039, 70.08860778808594, 69.70597076416016, 69.48533630371094, 68.90263366699219, 68.84700012207031, 68.53485107421875, 67.69471740722656, 66.90360260009766, 65.70887756347656, 65.6533203125, 64.92096710205078, 64.22373962402344, 63.748023986816406)

longitudes <- c(29.996711730957, 31.5819625854492, 27.7380523681641, 26.5001354217529, 26.6523590087891, 25.9216632843018, 22.900276184082, 23.2572174072266, 23.335693359375, 22.8744430541992, 23.0846500396729, 22.5654735565186, 21.4527740478516, 21.6638870239258, 21.0659694671631, 21.6765956878662, 21.4968719482422, 22.3399982452393, 22.2881927490234, 24.5395812988281, 25.4442329406738, 25.3037490844727, 24.6691665649414, 24.6891632080078, 24.1749992370605, 23.6847190856934, 24.0007610321045, 
23.5733299255371, 23.76513671875, 23.4308300018311, 23.6597900390625, 20.5809288024902, 21.3208312988281, 22.3983306884766, 23.9763870239258, 24.9349174499512, 25.7611083984375, 25.9593029022217, 26.4768047332764, 27.9106922149658, 29.1027774810791, 29.298469543457, 28.4355525970459, 28.8173580169678, 28.4598579406738, 30.0286102294922, 29.0751361846924, 30.1349277496338, 29.8188858032227, 29.6408309936523, 30.5773582458496, 29.996711730957)

poly <- sf::st_polygon(x = list(cbind(longitudes,latitudes)),dim = "XY")
```
 

```{r}
data <- sf::read_sf("./Automating-GIS-processes/source/exercises/exercise-2-ferdyndongo/Kruger_posts.shp")
data %>% sf::st_crs()
```
 
```{r}
data <- data %>% sf::st_transform(32735)
data %>% sf::st_crs()
```

```{r}
data %>% dplyr::group_by(userid) %>% dplyr::filter(userid==16301) %>%  dplyr::arrange(timestamp) %>% dplyr::group_split()
data

```
 

# Lesson 3: PIP (Point In Polygon) and intersect queries

Let’s create 2 point object and a polygon:

```{r}
p1 <- sf::st_point(x = c(24.952242, 60.1696017),dim = "XY")
p2 <- sf::st_point(x = c(24.976567, 60.1612500),dim = "XY")
poly <- sf::st_polygon(list(poly=rbind(c(24.950899, 60.169158), c(24.953492, 60.169158), c(24.953510, 60.170104), c(24.950958, 60.169990),c(24.950899, 60.169158))))
```
Let’s check if those points are within the polygon:

```{r}
poly %>% st_contains(p1)
```
 
```{r}
p1 %>% sf::st_within(poly)
```

```{r}
poly %>% st_contains(p2)
```

```{r}
p2 %>% sf::st_within(poly)
```

## Intersect
Let’s create two LineStrings:
```{r}
line_a <- st_linestring(rbind(c(0, 0), c(1, 1)))
line_b <- st_linestring(rbind(c(1, 1), c(0, 2)))
line_a %>% st_intersects(line_b)
```
Let’s see if they intersect
```{r}
# plot(line_a)
# plot(line_b,add=T)
ggplot()+ geom_sf(data = line_a)+geom_sf(data = line_b)
```
 
## Point in Polygon

* Reading KML-files
```{r}
polys <- sf::st_read("./Automating-GIS-processes/source/notebooks/L3/data/PKS_suuralue.kml")
southern <- sf::st_read("./Automating-GIS-processes/source/notebooks/L3/data/PKS_suuralue.kml") %>% filter(Name=='Eteläinen')
data <- sf::st_read("./Automating-GIS-processes/source/notebooks/L3/data/addresses.shp")
```
 
```{r}
ggplot()+geom_sf(data=polys,col="blue")+geom_sf(data=southern,col="red")+geom_sf(data=data)
```

```{r}
data %>% dplyr::filter(st_within(x = data,y = southern,sparse = FALSE))
```
```{r}
ggplot()+geom_sf(data=polys,col="blue")+geom_sf(data=southern,col="red")+geom_sf(data=data %>% dplyr::filter(st_within(x = data,y = southern,sparse = FALSE)))
```
 
```{r}
intersections <- sf::st_read("./Automating-GIS-processes/source/notebooks/L3/data/uusimaa_intersections.gpkg")
postcode_areas <- sf::st_read("./Automating-GIS-processes/source/notebooks/L3/data/uusimaa_postal_code_areas.gpkg")
```
 
```{r}
ggplot()+geom_sf(data=postcode_areas)+geom_sf(data=intersections,color="yellow")
```

```{r}
city_center_zip_area <- postcode_areas %>% dplyr::filter(posti_alue=='00100')
final_selection <- intersections %>% dplyr::filter(intersections %>% st_intersects(city_center_zip_area,sparse = FALSE))
```
 
```{r}
ggplot()+geom_sf(data=city_center_zip_area, color="red")+geom_sf(data=final_selection)
```

```{r}
# sf::st_join(x = intersections, y = postcode_areas) %>% group_by(posti_alue) %>% dplyr::count()
 
intersection_cnt <- sf::st_join(x = intersections, y = postcode_areas) %>% group_by(posti_alue) %>% dplyr::count() %>% dplyr::rename('intersection_cnt' = n)
# postcode_areas <- sf::st_intersection(x = intersections, y = postcode_areas) %>% group_by(posti_alue) %>% dplyr::count() %>% dplyr::rename('intersection_cnt' = n) %>% sf::st_drop_geometry() %>% dplyr::inner_join(postcode_areas)

# postcode_areas <- postcode_areas %>% dplyr::inner_join(sf::st_join(x = intersections, y = postcode_areas) %>% group_by(posti_alue) %>% dplyr::count() %>% dplyr::rename('intersection_cnt' = n) %>% sf::st_drop_geometry())
postcode_areas <- postcode_areas %>% dplyr::inner_join(intersection_cnt %>% st_drop_geometry())
# m2_to_km2_converter <- 1000000
postcode_areas$area <- postcode_areas %>% st_area() %>% units::set_units("km^2")
postcode_areas$intersection_density <- postcode_areas$intersection_cnt / postcode_areas$area
```
 
```{r}
plot(postcode_areas['intersection_density'])
```
 
## Spatial join o spatial filter

```{r}
wfs_url <- 'https://kartta.hsy.fi/geoserver/wfs'
httr::parse_url(wfs_url)
```
 
```{r}
url <- httr::parse_url(wfs_url)
url$query <- list(service = "WFS", version = "2.0.0", request = "GetCapabilities")
httr::build_url(url = url)
```

```{r}
url$query <- list(service = "WFS",
                  version = "2.0.0",
                  request = "GetFeature",
                  typeName='asuminen_ja_maankaytto:Vaestotietoruudukko_2018',
                  outputFormat='json'
                  )
 
httr::build_url(url = url)
```
 
```{r}
r <- httr::build_url(url = url)
pop <- read_sf(r)
pop %>% head()
```

```{r}
pop <- pop %>% dplyr::rename('pop18'=asukkaita) %>% dplyr::select(pop18)
pop %>% head()
```
 
```{r}
plot(pop)
```

## Join the layers
 

```{r}
addresses <- read_sf("./Automating-GIS-processes/source/notebooks/L3/data/addresses.shp")
addresses <- addresses %>% sf::st_transform(3879)
```
 
```{r}
addresses %>% st_join(pop,join = st_intersects, left=F) %>% dplyr::select(addr,pop18)
```
 

```{r}
addresses %>% st_join(pop,join = st_intersects, left=T) %>% dplyr::select(addr,pop18)
```

```{r}
addresses %>% st_join(pop,join = st_within, left=T) %>% dplyr::select(addr,pop18)
```

```{r}
addresses %>% st_join(pop,join = st_within, left=F) %>% dplyr::select(addr,pop18)
```
 

```{r}
ggplot()+geom_sf(data = pop,aes(fill=pop18)) +geom_sf(data=addresses,colour="red")
```
 
```{r}
join <- addresses %>% st_join(pop,join = st_within, left=F) %>% dplyr::select(addr,pop18)
```
 

```{r}
ggplot()+geom_sf(data = pop,aes(fill=pop18)) +geom_sf(data=join,colour="red")
```
 
```{r}
ows4R::WFSClient$new(url = wfs_url,serviceVersion = "2.0.0")
```

```{r}
client  <- ows4R::WFSClient$new(url = wfs_url,serviceVersion = "2.0.0")
```

```{r}
client$getFeatureTypes(pretty = TRUE)
```

```{r}
client$getCapabilities()$findFeatureTypeByName("asuminen_ja_maankaytto:1000m_verkostobufferi")$getDescription() %>% map_chr(function(x){x$getName()})
```
 
```{r}
client$getCapabilities()$getOperationsMetadata()$getOperations() %>% 
     purrr::map_chr(function(x){
         x$getName()
     })
```

```{r}
client$getCapabilities()$getOperationsMetadata()$getOperations() %>% 
     purrr::map(function(x){
         x$getParameters()
     })
```
 
```{r}
client$getCapabilities()$ getFeatureTypes() %>%  
  map(function(x){x$getBoundingBox()})
```
 
```{r}
client$getCapabilities()$ getFeatureTypes() %>%  
  map_chr(function(x){x$getAbstract()})
```
 
```{r}
wfs_url <- 'https://kartta.hsy.fi/geoserver/wfs'
httr::parse_url(wfs_url)
```
 
```{python python.reticulate = FALSE}
import requests
from pyproj import CRS
import geopandas as gpd
import geojson
url = 'https://kartta.hsy.fi/geoserver/wfs'
# Specify parameters (read data in json format). 
# Available feature types in this particular data source: 
# http://geo.stat.fi/geoserver/vaestoruutu/wfs?service=wfs&version=2.0.0&request=describeFeatureType

params = dict(service='WFS', 
              version='2.0.0', 
              request='GetFeature', 
              typeName='asuminen_ja_maankaytto:Vaestotietoruudukko_2018', 
              outputFormat='json')
# Fetch data from WFS using requests
r = requests.get(url, params=params)
# Create GeoDataFrame from geojson
pop = gpd.GeoDataFrame.from_features(geojson.loads(r.content))
```

```{python python.reticulate = FALSE}
pop.head()
```


## Nearest neighbour analysis
One commonly used GIS task is to be able to find the nearest neighbour for an object or a set of objects. For instance, you might have a single Point object representing your home location, and then another set of locations representing e.g. public transport stops. Then, quite typical question is “which of the stops is closest one to my home?” This is a typical nearest neighbour analysis, where the aim is to find the closest geometry to another geometry.

```{r}
orig <- sf::st_point(x = c(1, 1.67), dim = "XY")
dest1 <- sf::st_point(x = c(0, 1.45), dim = "XY")
dest2 <- sf::st_point(x = c(2, 2), dim = "XY")
dest3 <- sf::st_point(x = c(0, 2.5), dim = "XY")
destinations <- sf::st_multipoint(c(dest1,dest2,dest3))
# sf::st_nearest_feature(x = orig,y = destinations) # the index of the nearest feature
```

```{r}
nearest_geoms <- sf::st_nearest_points(orig, destinations)
nearest_geoms
```

```{r}
ggplot()+geom_sf(data=orig,colour="blue")+geom_sf(data=destinations,colour="red")+geom_sf(data=nearest_geoms,colour="green")
```
 
```{r}
destinations[sf::st_nearest_feature(orig,destinations),]
```
 
```{r}
df2 <- read_sf("./Automating-GIS-processes/source/notebooks/L3/data/addresses.shp")
df1 <- read_sf("./Automating-GIS-processes/source/notebooks/L3/data/PKS_suuralue.kml")
```

```{r}
df1$centroid <-  st_geometry(df1) %>% st_centroid()
df1 <- st_set_geometry(df1,'centroid') # activate centroid as active geometry
df1 %>% st_nearest_points(st_union(df2))
```

```{r}
df1 %>% st_nearest_points(st_union(df2))
```
 
```{r}
df1 %>% st_nearest_feature(df2)
```

```{r}
ggplot()+geom_sf(data=df1,colour="green")+geom_sf(data=df2,colour="red")+geom_sf(data=df1 %>% st_nearest_points(st_union(df2)))
```

```{r}
st_nearest_feature(df1[2,],df2)
```

```{r}
st_nearest_points(df1[2,],st_union(df2))
```

```{r}
st_nearest_points(df1[2,],df2)
```

 
```{r}
ggplot()+geom_sf(data=df1[1,],colour="green")+geom_sf(data=df2,colour="red")+geom_sf(data=df1[1,] %>% st_nearest_points(st_union(df2)),colour="blue")
```
 

```{r}
df2 %>% dplyr::slice(df1 %>% st_nearest_feature(df2)) %>% dplyr::select(id, addr)
```
 

```{r}
df1$nearest_point <- df1 %>% sf::st_nearest_points(sf::st_union(df2))
df1
```
 

```{r}
buildings <- read_sf(unzip("./Automating-GIS-processes/source/notebooks/L3/data/building_points_helsinki.zip"))
stops <- read_sf("./Automating-GIS-processes/source/notebooks/L3/data/pt_stops_helsinki.gpkg")
closest_stops <- buildings %>% sf::st_nearest_points(st_union(stops))
```
```{r}
par(mfrow=c(1,3))
buildings %>% ggplot()+geom_sf()+ggtitle("Buildings")
stops %>% ggplot()+geom_sf()+ggtitle("Stops")
closest_stops %>% ggplot()+geom_sf()+ggtitle("Closest Stops")
```
 
```{r}
buildings_closest_stops <- buildings %>% dplyr::bind_cols(stops %>% dplyr::slice(buildings %>% st_nearest_feature(stops)))
buildings_closest_stops <- buildings_closest_stops %>% dplyr::rename('building_centroid'=geom...2,'closest_stop'=geom...7)
buildings_closest_stops <- buildings_closest_stops %>% st_as_sf()
```
 

```{r}
buildings_closest_stops$distance <- st_geometry(buildings_closest_stops) %>% st_distance(st_geometry(st_set_geometry(buildings_closest_stops,'closest_stop')),by_element = T)
buildings_closest_stops$linestring <- closest_stops
```
 

```{r fig.width=10}
ggplot()+geom_sf(data=st_set_geometry(buildings_closest_stops,'closest_stop'),colour="red")+geom_sf(data=buildings_closest_stops,colour="yellow")+geom_sf(data=st_set_geometry(buildings_closest_stops,'linestring'),colour="green")
 

```
 

## Exercise 3 of Autogis process with python (with R code)
 

```{r}
geodata <- sf::st_read('./Automating-GIS-processes/source/exercises/exercise-3-ferdyndongo/shopping_centers.shp')
geodata$buffer <- geodata %>% sf::st_buffer(dist = 1500) %>% sf::st_geometry()
```

```{r}
geodata$area <- geodata %>% st_set_geometry(value = 'buffer') %>% st_area() %>% units::set_units(value = 'km2')
geodata <- geodata %>% st_set_geometry(value = 'buffer')
```

```{r}
wfs_url <- 'https://kartta.hsy.fi/geoserver/wfs'
url <- httr::parse_url(wfs_url)
url$query <- list(service = "WFS",
                   version = "2.0.0",
                   request = "GetFeature",
                   typeName='asuminen_ja_maankaytto:Vaestotietoruudukko_2018',
                   outputFormat='json'
                  )
 
r <- httr::build_url(url = url)
pop <- read_sf(r)
pop <- pop %>% dplyr::rename('pop18'=asukkaita) %>% dplyr::select(pop18)
```

```{r}
ggplot()+geom_sf(data=geodata,colour="lightgreen")+geom_sf(data=st_set_geometry(x = geodata,value = 'geometry'),colour="red")
```

```{r}
ggplot()+geom_sf(data=geodata,colour="red")+geom_sf(data=st_set_geometry(x = geodata,value = 'geometry'),colour="red")+geom_sf(data=pop)
```
 

```{r}
geodata %>% st_join(y = pop,join = st_contains) %>% dplyr::group_by(id) %>% dplyr::summarise(pop=sum(pop18))
```
 

```{r}
ggplot()+geom_sf(data=pop %>% st_join(y = geodata, join = st_within,left=F))+ geom_sf(data=geodata %>% st_join(y = pop,join = st_contains),colour="red")
 
```
 
```{r include=FALSE}
geodata <- geodata %>%  dplyr::rename( geometry=buffer, shopping_center = geometry) %>% st_set_geometry(value = 'geometry')
 
```
 

 
## Geometric operation: overlay analysis
 
```{r}

hel <- st_read('./Automating-GIS-processes/source/notebooks/L4/data/Helsinki_borders.shp')
 
grid <- st_read('./Automating-GIS-processes/source/notebooks/L4/data/TravelTimes_to_5975375_RailwayStation.shp')
```
 
```{r echo=FALSE, include=TRUE, warning=FALSE}
ggplot()+geom_sf(data=grid)+geom_sf(data=hel)
```

```{r}
# grid %>% dplyr::filter(grid %>% st_intersects(y = hel,sparse = F))
intersection <- grid %>% st_join(y = hel, join = st_intersects,left=F)
```
 
```{r}
ggplot()+geom_sf(data=intersection)
```
 

```{r echo=FALSE}
st_write(obj = intersection,dsn = 'intersection.geojson')
```

* Aggregating data
 
```{r}
dissolved <- intersection %>% dplyr::group_by(car_r_t) %>% summarise()
selection <- dissolved %>% dplyr::filter(car_r_t==15)
```
 

```{r}
ggplot()+geom_sf(data=dissolved)+geom_sf(data=selection,colour="red")
```
 
* Simplify geometry
 

```{r}
amazon_river <- st_read('./Automating-GIS-processes/source/notebooks/L4/data/Amazon_river.shp')
```
 

```{r}
ggplot()+geom_sf(data=amazon_river)
```
 

```{r}
ggplot()+geom_sf(data = st_simplify(amazon_river))
```
 

## Data reclassification
 
```{r}
acc <- st_read('./Automating-GIS-processes/source/notebooks/L4/data/TravelTimes_to_5975375_RailwayStation_Helsinki.geojson')
acc <- acc %>% dplyr::select(pt_r_tt,walk_d) %>% dplyr::filter(pt_r_tt >= 0)
```
 

```{r}
tm_shape(acc)+tmap::tm_fill('pt_r_tt')
```
 

```{r}
ggplot(data=acc) + geom_sf(aes(fill=cut(pt_r_tt,classInt::classIntervals(acc$pt_r_tt,n = 9,style = 'quantile')$brks)))+scale_fill_brewer(palette = 'RdYlBu',name='times -> center')
```
 

```{r}
ggplot() + geom_sf(data=acc,aes(fill=pt_r_tt %>% cut_to_classes(n = 9,style = 'jenks'))) + scale_fill_brewer(palette= 'RdYlBu',name='travel times')
```
 

```{r}
ggplot(data=acc) + geom_sf(aes(fill=cut(walk_d,classInt::classIntervals(acc$walk_d,n = 9,style = 'quantile')$brks)))+scale_fill_brewer(palette = 'RdYlBu',name='walking distance')
```
 

```{r echo=FALSE}
ggplot() + geom_sf(data=acc,aes(fill= walk_d %>% cut_to_classes(n = 9,style = 'jenks'))) + scale_fill_brewer(palette= 'RdYlBu',name='walking distances')
```
 

* with fisher
 
```{r}
ggplot() + geom_sf(data=acc,aes(fill= walk_d %>% cut_to_classes(n = 9,style = 'fisher'))) + scale_fill_brewer(palette= 'RdYlBu',name='walking distances')
```
 
* with tmap packages
```{r}
tm_shape(acc)+tmap::tm_fill('walk_d')
```
```{r}
tm_shape(acc)+tmap::tm_fill('walk_d',midpoint = NA)
```
 
 
```{r echo=FALSE}
acc$nb_pt_r_tt <-  cut_to_classes(x = acc$pt_r_tt, n = 9,style = 'jenks')
```
 
```{r echo=FALSE}
ggplot(data = acc) + geom_sf(aes(fill=nb_pt_r_tt))
```
 
# AutoGis Processing: exercise 4
 
```{r}
grid <- st_read('./Automating-GIS-processes/source/exercises/exercise-4-ferdyndongo/data/MetropAccess_YKR_grid_EurefFIN.shp')
TravelTimes_to_Itis <- read.csv("./Automating-GIS-processes/source/exercises/exercise-4-ferdyndongo/data/TravelTimes_to_5944003_Itis.txt", sep=";") %>% dplyr::select(pt_r_t, car_r_t, from_id, to_id)
geodata <- grid %>% dplyr::inner_join(y = TravelTimes_to_Itis,by=c('YKR_ID'='from_id'))
geodata <- geodata %>% dplyr::filter( pt_r_t != -1 )
```
 

```{r echo=FALSE}
geodata <- geodata %>% dplyr::mutate(pt_r_t_cl=findInterval(x = geodata$pt_r_t,vec = c(5, 10, 15, 20, 25, 30, 35, 40, 45, 50, 55, 60)))
geodata <- geodata %>% dplyr::mutate(car_r_t_cl=findInterval(x = geodata$car_r_t,vec = c(5, 10, 15, 20, 25, 30, 35, 40, 45, 50, 55, 60)))
```
 
```{r echo=FALSE, warning=FALSE, fig.asp=0.5, fig.width=5,fig.align='center'}
par(mfrow=c(1,2))
ggplot(data = geodata) + geom_sf(aes(fill=pt_r_t_cl)) + scale_color_brewer(palette= 'RdYlBu')
ggplot(data = geodata) + geom_sf(aes(fill=car_r_t_cl)) + scale_color_brewer(palette= 'RdYlBu')
```

```{r echo=FALSE}
dir_name <- './Automating-GIS-processes/source/exercises/exercise-4-ferdyndongo/data/'
files <- list.files(path = dir_name,pattern = '*.txt')
for (f in files){
    center_name <- str_split(string = f,pattern = '_')[[1]][4] %>% str_replace(pattern = '.txt',replacement = '')
    data <- read.csv(file =paste(dir_name,f, sep = ''),sep = ';') %>% dplyr::select(from_id,pt_r_t)
    names(data)[names(data)=='pt_r_t'] <- paste('pt_r_t',center_name,sep='_')
    if (! exists('travel_time')){
      travel_time <- data
    }else{
      travel_time <- travel_time %>% dplyr::inner_join(data,by='from_id')
    }
    rm(data)
}
# travel_time %>% dplyr::rowwise() %>% dplyr::mutate( min_t=min(pt_r_t_Jumbo, pt_r_t_Dixi, pt_r_t_Myyrmanni, pt_r_t_Itis, pt_r_t_Forum, pt_r_t_IsoOmena, pt_r_t_Ruoholahti))
# data %>% dplyr::rowwise() %>% dplyr::mutate( min_t=min(pt_r_t_Jumbo, pt_r_t_Dixi, pt_r_t_Myyrmanni, pt_r_t_Itis, pt_r_t_Forum, pt_r_t_IsoOmena, pt_r_t_Ruoholahti))
travel_time <- travel_time[which(!travel_time==-1),]
travel_time$min_t <- apply(travel_time, 1, min)
travel_time$dominant_service <- apply(travel_time, 1, which.min)
data <- grid %>% dplyr::inner_join(travel_time, by=c('YKR_ID'='from_id'))
```
 
```{r echo=FALSE}
par(mfrow=c(1,2))
ggplot(data) + geom_sf(aes(fill=dominant_service))+scale_color_discrete(palette= 'RdYlBu',name='dominant service')
ggplot(data) + geom_sf(aes(fill=min_t))+scale_color_discrete(palette= 'RdYlBu',name='travel time')
```
 
```{r}
dir_name <- './Automating-GIS-processes/source/exercises/exercise-4-ferdyndongo/data/'
files <- list.files(path = dir_name,pattern = '*.txt')
for (f in files){
    center_name <- str_split(string = f,pattern = '_')[[1]][4] %>% str_replace(pattern = '.txt',replacement = '')
    data <- read.csv(file =paste(dir_name,f, sep = ''),sep = ';') %>% dplyr::select(from_id,pt_r_t)
    names(data)[which(names(data)=='from_id')] <- 'YKR_ID'
    names(data)[names(data)=='pt_r_t'] <- paste('pt_r_t',center_name,sep='_')
    grid <-grid %>% dplyr::inner_join(data,by='YKR_ID')
}
grid <- grid %>% dplyr::filter(pt_r_t_Jumbo != -1)
# grid <- grid %>% dplyr::rowwise() %>% dplyr::mutate( min_t=min(pt_r_t_Jumbo, pt_r_t_Dixi, pt_r_t_Myyrmanni, pt_r_t_Itis, pt_r_t_Forum, pt_r_t_IsoOmena, pt_r_t_Ruoholahti))
grid <- grid %>% dplyr::bind_cols(grid %>% dplyr::rowwise() %>% transmute( min_t=min(pt_r_t_Jumbo, pt_r_t_Dixi, pt_r_t_Myyrmanni, pt_r_t_Itis, pt_r_t_Forum, pt_r_t_IsoOmena, pt_r_t_Ruoholahti)))
 
grid$dominant_service <- colnames(grid[,c('pt_r_t_Jumbo', 'pt_r_t_Dixi', 'pt_r_t_Myyrmanni', 'pt_r_t_Itis', 'pt_r_t_Forum', 'pt_r_t_IsoOmena', 'pt_r_t_Ruoholahti')])[apply(grid[,c('pt_r_t_Jumbo', 'pt_r_t_Dixi', 'pt_r_t_Myyrmanni', 'pt_r_t_Itis', 'pt_r_t_Forum', 'pt_r_t_IsoOmena', 'pt_r_t_Ruoholahti')],1,which.min)]
 
grid$dominant_service_idx <- apply(grid[,c('pt_r_t_Jumbo', 'pt_r_t_Dixi', 'pt_r_t_Myyrmanni', 'pt_r_t_Itis', 'pt_r_t_Forum', 'pt_r_t_IsoOmena', 'pt_r_t_Ruoholahti')],1,which.min)
 
```
 

```{r echo=FALSE}
par(mfrow=c(1,2))
ggplot(grid) + geom_sf(aes(fill=dominant_service))+scale_color_discrete(palette= 'RdYlBu',name='dominant service')
 
ggplot(grid) + geom_sf(aes(fill=min_t))+scale_color_discrete(palette= 'RdYlBu',name='travel time')
```
 

## AutoGis processing with R and QGis (lesson 7)
 

```{r}
wfs_url <- 'http://kartta.hel.fi/ws/geoserver/avoindata/wfs'
client <- ows4R::WFSClient$new(url = wfs_url,serviceVersion = "2.0.0")
```
 
```{r}
client$getFeatureTypes(pretty = TRUE) %>% head()
```
 
```{r}
client$getCapabilities()
```
 
```{r}
client$getCapabilities()$findFeatureTypeByName('Seutukartta_liikenne_metro_rata')
```

```{r}
client$getCapabilities()$findFeatureTypeByName('Seutukartta_liikenne_metro_rata')$getDescription(pretty=FALSE)
```
 
```{r}
client$getCapabilities()$findFeatureTypeByName('Seutukartta_liikenne_metro_rata')$getDescription(pretty=FALSE) %>% map_chr(function(x){x$getName()})
```
 
```{r}
client$getCapabilities()$getOperationsMetadata()
```

```{r}
client$getCapabilities()$getOperationsMetadata()$getOperations()
```

```{r}
client$getCapabilities()$getOperationsMetadata()$getOperations() %>% map_chr(function(x){x$getName()})
```
 
```{r}
client$getCapabilities()$getOperationsMetadata()$getOperations() %>% purrr::map(function(x){x$getParameters()})
```
 
```{r}
url <- httr::parse_url(wfs_url)
url$query <- list(service = "WFS",
                   version = "2.0.0",
                   request = "GetFeature",
                   typeName='Seutukartta_liikenne_metro_rata',
                   outputFormat='json'
                  )
r <- httr::build_url(url = url)
r
```
 
```{r}
metro_data <- sf::read_sf(r)
```
 
```{r}
ggplot()+geom_sf(data = metro_data)
```
 